Interfacing logic programming and Coq

### 2020-10-28 papers/types2020/pbt.tex
commit before Wed's meeting

Let us consider a  relational specification of ordered list, as well as insertion in such lists:
\begin{lstlisting}
Inductive insert (x:nat) : list nat -> list nat -> Prop :=
| i_n : is_nat x -> insert x [] [x]
| i_s : forall y: nat, forall ys, x <= y -> insert x (y :: ys) (x :: y :: ys)
| i_c : forall y: nat, forall ys rs, x > y -> insert x ys rs -> insert x (y :: ys) (x :: rs).

Inductive ordered : list nat -> Prop :=
| o_n : ordered []
| o_s : forall x : nat, ordered [x]
| o_c : forall (x y : nat), forall xs,  ordered xs  -> x <= y -> ordered (x:: y :: xs).     
\end{lstlisting}
%
A property we may wish to check before embarking in a formal proof is whether insertion preserves ordered-ness. Since we did slip in an error, our tactic reports a counter-example, e.g.,
\verb|(* Proof Term: [0, [0; 1; 0]] *)|:
\begin{lstlisting}
Conjecture ins_ord:  forall x xs rs, ordered xs -> insert x xs rs -> ordered rs.
intros x xs rs  Ho Hi.
elpi dep_pbt height 5 (Ho /\ Hi ) (x) (xs).
Abort.
\end{lstlisting}
In this query the tactic tests the hypotheses \lsti{Ho} and \lsti{Hi}
against data \lsti{x,xs} generated exhaustively up to proofs bounded by
at most $5$ unfoldings of the library \lsti{Inductive} definition of
\lsti{nat} and \lsti{list}.

>>>>

Let us consider a  relational specification of ordered list:
\begin{lstlisting}
Inductive ordered : list nat -> Prop :=
| o_n : ordered []
| o_s : forall x : nat, ordered [x]
| o_c : forall (x y : nat), forall xs,  ordered xs  -> x <= y -> ordered (x:: y :: xs).     
\end{lstlisting}
%
A property we may wish to check before embarking in a formal proof is whether insertion preserves ordered-ness. Since we did slip in an error, our tactic reports a counter-example, e.g.,
\verb|(* Proof Term: [0, [0; 1; 0]] *)|:
\begin{lstlisting}
Conjecture ins_ord:  forall x xs rs, ordered xs -> insert x xs rs -> ordered rs.
intros x xs rs  Ho Hi.
elpi dep_pbt height 5 (Ho /\ Hi ) (x) (xs).
Abort.
\end{lstlisting}
In this query the tactic tests the hypotheses \lsti{Ho} and \lsti{Hi}
against data \lsti{x,xs} generated exhaustively up to proofs bounded by
at most $5$ unfoldings of the library \lsti{Inductive} definition of
\lsti{nat} and \lsti{list}.
